---
title: "Milestone 9: Attribution Module Memo"
author: "Kayla Cortez, Matthew Melgar, Jamie Min, Mehrnoosh Solgi, Roxana Rodriguez"
date: today
format:
  html:
   theme: cosmo
   toc: true
   toc-depth: 4
   number-sections: false
   code-line-numbers: true
   code-link: true
   code-fold: false
   embed-resources: true
editor: visual
execute:
  freeze: auto
  warning: false
---

**Repo Link:** <https://github.com/roxanar1/IBM6530-Milestone-9.git>

## Deliverable 1 — Path Dataset Definition

### Unit of Analysis

The unit of analysis for NCMA SGV GA4 website data consists of a user-level conversion path that has been structured from session-level data. For example, we used identifiers such as user_pseudo_id and ga_session_id to combine multiple sessions and create the full user journey from the website. Each path contains multiple sessions that belong to a single user within a specific timeframe, which helps us identify multiple user behaviors across channels and to understand how users interact with the website before completing a desired action, such as membership purchase or newsletter sign-up. 

### What counts as a touchpoint

A touchpoint isa session-level interaction that is usually linked to a type of traffic source or medium. Each session contains one touchpoint by traffic channel, such as organic search, direct, or referral, and sources such as Google, direct, or none. The data also shows additional dimensions such as device category and geographic location. From the data it can be seen that each row is a single touchpoint containing engagement metrics such as engaged session and average engagement time. 

### Lookback Window

For the dataset, the lookback window can be defined within a 30-day timeframe, meaning that all sessions within this period are considered in the user’s conversion path. A 30-day timeframe is appropriate because it captures all interactions across the different channels and reflects the user behavior, such as how they engage with this site, since potential customers for NCMA SGV may like to explore the membership information before purchasing. This lookback window also provides enough time to capture user interactions over time. 

### How you handle direct/none, missing UTMs, cross-device limitations

From previous Mart views in BigQuery, direct traffic is often defined as "direct" and "none," and it’s included as a valid touchpoint. Also, if a prior session has a known source of traffic, this may be interpreted as continued engagement rather than a new acquisition channel, which may avoid over-attribution. A limitation the NCMA SGV website has is that it has missing UTM parameters, so when the traffic source is known, GA4 would define it as “Google” or with a specific name such as "link.edgepilot.com." However, when the traffic origins are not known, GA4 categorizes the website traffic that should have a UTM parameter as “Unknown.” Another limitation of the GA4 dataset is that because it relies on user_pseudo_id, users that interact with the website from multiple devices may be treated as separate users and this may fragment the user conversion path. 

### How you define “conversion” 

For the NCMA SGV website, a conversion should be defined as a session in which a user successfully purchased a membership subscription or event registration, as this aligns with the primary goal of driving membership growth and event participation. Some user conversions can include form submissions to classify a user as interested in the memberships, clicks on registration buttons or simply general engagement such as views on event pages. All these website user behaviors can give insights on the level of user intent to analyze how much interest there is in the memberships and events but also on overall website performance. 

## Deliverable 2 — Method Rationale

We selected position-based (U-shaped) attribution as our primary method, with last-click as a comparison baseline.

Explain why your method matches:

Position-based attribution splits credit across the conversion path by role rather than sequence. The first and last touchpoints each receive 40% of the credit, with the remaining 20% distributed evenly across any middle interactions. This fits our client context for two reasons.

First, NCMA SGV is a membership-driven nonprofit where awareness genuinely matters. A user who first discovers the organization through organic search behaves differently from one who comes back directly after receiving an email. Giving the first touch meaningful weight captures that initial acquisition moment, which last-click attribution would erase entirely.

Second, the final touchpoint before a membership purchase also matters, it's what pushed the user to convert. Giving it equal weight to the first touch reflects that both ends of the journey are strategically important for a client deciding where to invest in outreach versus retention.

Why last-click as a baseline? Last-click is what most GA4 reports default to, and it's what the client likely already sees in their dashboard. Running both models lets us flag the gap: channels that look weak under last-click but score higher under position-based are probably doing awareness or mid-funnel work that would be quietly cut if the client relied on GA4 defaults alone. That comparison is where the actionable insight lives.

What this method can and cannot do: Position-based attribution applied to aggregated data is a reasonable approximation, not a precise causal model. We're inferring channel roles from volume and conversion patterns rather than tracking individual journeys. The model assumes channels behave consistently across the lookback window, which may not hold if campaign activity shifted significantly between February and March. These constraints go directly into our limitations section.

## Deliverable 3 — Results, Implications, Limitations, Next Tests (2–4 pages)

### Attribution Module

```{r}
library(arrow)
library(dplyr)
library(ggplot2)

data_path <- "channel_daily_kpis_v1"

ds <- open_dataset("channel_daily_kpis_v1")

ds
```

### Ranked channels by attributed value

```{r}
channel_rank <- ds |>
  group_by(channel) |>
  summarise(
    total_sessions = sum(sessions, na.rm = TRUE),
    total_purchases = sum(purchases, na.rm = TRUE),
    total_revenue = sum(revenue, na.rm = TRUE)
  ) |>
  arrange(desc(total_purchases)) |>
  collect()

channel_rank
```

Channels were ranked based on total purchases as a proxy for attributed value. The results show that while some channels generate higher traffic, they do not necessarily produce the highest number of purchases, indicating differences in channel effectiveness.

### Campaign-level (use source + medium)

```{r}
campaign_rank <- ds |>
  group_by(source, medium) |>
  summarise(
    total_sessions = sum(sessions, na.rm = TRUE),
    total_purchases = sum(purchases, na.rm = TRUE)
  ) |>
  arrange(desc(total_purchases)) |>
  collect()

campaign_rank
```

### Compare to last-click

#### **conceptual**

This analysis approximates attribution using aggregated metrics rather than a true multi-touch attribution model. Compared to a traditional last-click approach, which assigns full credit to the final interaction, this method highlights broader channel contributions through engagement and session volume. As a result, some channels that may appear less valuable under last-click attribution (due to fewer direct conversions) may still play an important role in earlier stages of the user journey.

**Recommendation 1 — Optimize high-conversion channels** Channels with higher purchase-to-session ratios should be prioritized for budget allocation, as they demonstrate stronger conversion efficiency compared to high-traffic but low-converting channels.

**Recommendation 2 — Investigate traffic spikes** The presence of spikes in session activity suggests that certain campaigns or events drive short-term increases in traffic. Identifying and replicating these drivers could help stabilize and grow user engagement over time.

**Recommendation 3 — Improve low-performing segments** Channels or device categories with high traffic but low engagement or conversion rates should be further analyzed to identify potential issues such as poor targeting, weak messaging, or suboptimal user experience.

Additional tracking or attribution modeling should be implemented to better understand multi-touch customer journeys and improve decision-making beyond last-click metrics.

```{r}
channel_rank
```

## Deliverable 4 — Limitations + Next Tests

Be explicit about:

-   **Tracking gaps**

-   When it comes to analyzing GA4 data metrics from sessions, purchases, revenue which is grouped by channel and campaign. This means that the accuracy depends consistent tracking and tagging. If we don't use UTM tracking parameters properly it can cause traffic to misclassified. This will impact how channels are ranked and interpreted

-   **Missing touchpoints (walled gardens, offline)**

    -   GA4 may not capture all interactions. Activity outside the website such as LinkedIn and other social media channels may not provide the full picture, it will provide us partial data. For example, channel source and where are users coming from when they visit the website, but not the early-stage interactions. This means that the analysis that we are doing is only what is visible to us through the website.

-   **Low volume uncertainty**

    -   If NCMA website has low website traffic or purchase counts the results may become less reliable. Small changes in the data can impact how the results will be interpreted, such as if we have a campaign it may influence our results to be seen not effective or more than effective.

-   **Suggested next tests (measurement improvements, experiments, etc.)**

    -   Suggested improvements:

        -   Implement consistent UTM naming conventions across future campaigns such as source, medium, and campaign. With these in mind it will be important to audit tagging regularly. As well as enable or look into in being able to multi-touchpoints such as Markov Model. This will help capture channels accross the user's journey.